On the presentation of orthogonal groups over polynomial rings

### 2019-01-16 dl-paper.tex
More text added

\theoremstyle{definition}
\newtheorem{df}[lemma]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example} \Crefname{example}{Example}{Examples}
\newtheorem{rem}[lemma]{Remark}
\newtheorem{cond}[lemma]{Condition}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\GW}{GW}
\DeclareMathOperator{\colim}{colim}
\newcommand{\XX}{\mathcal{X}}           % (subset of) Steinberg generators
\newcommand{\RR}[1]{\mathcal{R}_{#1}}   % Steinberg relations

@@

The aim of this section is to briefly remind the reader definitions of relative Steinberg groups and their basic properties. We refer the reader to~\cite[Section~3]{S15} for a more detailed exposition.
%TODO: Define the category of pairs (so that all constructions below become functorial wrt it).

...

@@

\begin{prop} The group $\K_3(\Phi, R)$ is isomorphic to $\HH_3(\St(\Phi, R), \ZZ)$. In particular, it does not depend on the isogeny form of $\G(\Phi, R)$ used to define it. Moreover, for a ring epimorphism $R \to R/I$ there is a functorial exact sequence
\[ \xymatrix{\K_3(\Phi, R) \ar[r] & \K_3(\Phi, R/I) \ar@{->>}[r] & C(\Phi, R, I).} \]
\end{prop}
\begin{proof}
 %TODO: First part.

@@

\section{Proof of the main result}
\subsection{Overview of the proof}
\subsection{First injectivity theorem}
For a commutative local ring $A$ with the maximal ideal $M$ we denote by $B = B_{A, M}$ the subring $A[X^{-1}] + M[X]$ of the ring $R = A[X, X^{-1}]$.
It is clear that $I := M[X, X^{-1}]$ can be considered as an ideal of both $B$ and $R$.
\begin{lemma}\label{lem:lemma32} Let $\Phi$ be a simply-laced root system of rank $\geq 3$.
Consider the following commutative square of canonical maps.
\[ \xymatrix{
    \St(\Phi, B, I) \ar[r] \ar[d] & \St(\Phi, B) \ar[d] \\
    \St(\Phi, R, I) \ar[r] \ar@{-->}[ur]_\mu & \St(\Phi, R)
   } \]
Then there exists an arrow $\mu$ which makes the diagram commute.   
\end{lemma} 
\begin{proof}
 Notice that $R$ is isomorphic to the principal localisation of $B$ at $X$
  and that $I$ is uniquely $X$-divisible in the sense of~\cite[\S~4]{LS17}.
 Thus, in the special case $\Phi = \rA_3$ the assertion of the lemma follows from~\cite[Theorem~3]{LS17}.

@@

The following lemma summarizes facts about $\K_2(\Phi, -)$ in the field case.
\begin{lemma}
 Let $F$ be a field of characteristic $\neq 2$. Then for $\ell \geq 7$ the canonical map $\K_3(\rD_\ell, F[X]) \to \K_3(\rD_\ell, F[X^{\pm 1}])$ is surjective.
\end{lemma}
\begin{proof}
 We have defined $\K_3(\rD_\ell, R)$ as 
\end{proof}

\begin{theorem} Suppose $A$ is a local ring of characteristic $\neq 2$, then for $\ell \geq 7$
  the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective.
\end{theorem}
\begin{proof}

@@

\begin{equation*} \xymatrix{
 C(\Phi, B, I) \ar[r] \ar[d] & \St(\Phi, B, I) \ar[r] \ar[d] & \St(\Phi, B) \ar[r] \ar[d] & \St(\Phi, k[t]) \ar[d] \\
 C(\Phi, R, I) \ar[r]        & \St(\Phi, R, I) \ar[r] \ar@{-->}[ur]& \St(\Phi, R) \ar[r]        & \St(\Phi, k[t, t^{-1}])
}\end{equation*}

>>>>

\theoremstyle{definition}
\newtheorem{df}[lemma]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example} \Crefname{example}{Example}{Examples}
\newtheorem{rem}[lemma]{Remark}
\newtheorem{cond}[lemma]{Condition}
\DeclareMathOperator{\Ker}{Ker}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\GW}{GW}
\DeclareMathOperator{\colim}{colim}
\newcommand{\XX}{\mathcal{X}}           % (subset of) Steinberg generators
\newcommand{\RR}[1]{\mathcal{R}_{#1}}   % Steinberg relations

@@

The aim of this section is to briefly remind the reader definitions of relative Steinberg groups and their basic properties. We refer the reader to~\cite[Section~3]{S15} for a more detailed exposition.
%TODO: Define the category of pairs (so that all constructions below become functorial wrt it).
...

@@

\begin{prop}\label{characterization} The group $\K_3(\Phi, R)$ is isomorphic to $\HH_3(\St(\Phi, R), \ZZ)$. In particular, it does not depend on the isogeny form of $\G(\Phi, R)$ used to define it. Moreover, for a ring epimorphism $R \to R/I$ there is a functorial exact sequence
\[ \xymatrix{\K_3(\Phi, R) \ar[r] & \K_3(\Phi, R/I) \ar@{->>}[r] & C(\Phi, R, I).} \]
\end{prop}
\begin{proof}
 %TODO: First part.

@@

\section{Proof of the main result}
\subsection{Overview of the proof}
\subsection{First injectivity theorem}
For a commutative local ring $A$ with the maximal ideal $M$ and residue field $k$
we denote by $B = B_{A, M}$ the subring $A[X^{-1}] + M[X]$ of the ring $R = A[X, X^{-1}]$.
It is clear that $I := M[X, X^{-1}]$ can be considered as an ideal of both $B$ and $R$.
\begin{lemma}\label{lem:lemma32} Let $\Phi$ be a simply-laced root system of rank $\geq 3$.
Consider the following commutative square of canonical maps.
\[ \xymatrix{
    \St(\Phi, B, I) \ar[r] \ar[d] & \St(\Phi, B) \ar[d] \\
    \St(\Phi, R, I) \ar[r] \ar@{-->}[ur] & \St(\Phi, R)
   } \]
Then there exists a diagonal arrow which makes the diagram commute.   
\end{lemma} 
\begin{proof}
 Notice that $R$ is isomorphic to the principal localisation of $B$ at $X$
  and that $I$ is uniquely $X$-divisible in the sense of~\cite[\S~4]{LS17}.
 Thus, in the special case $\Phi = \rA_3$ the assertion of the lemma follows from~\cite[Theorem~3]{LS17}.

@@

The following lemma summarizes facts about $\K_2(\Phi, -)$ in the field case.
\begin{lemma} \label{lem:field-case}
 Under certain assumptions the map $\St(\Phi, k[t]) \to \St(\Phi, k[t, t^{-1}])$ is injective.
\end{lemma}
\begin{proof}
 %TODO: Add references on Rehmann and Hurrelbrink...
\end{proof}

\begin{prop}
 Suppose that the canonical map $C(\Phi, B, I) \to C(\Phi, R, I)$ is surjective,
  and that Tulenbaev lifting property holds for $\St(\Phi, R)$. 
 Then the canonical map of Steinberg groups $\St(\Phi, B) \to \St(\Phi, R)$ is injective.
\end{prop}
\begin{proof}
 Consider the following commutative diagram with exact rows: 
\begin{equation*} \xymatrix{
 C(\Phi, B, I) \ar[r]^{j_B} \ar@{->>}[d]_{f} & \St(\Phi, B, I) \ar[r]^{\mu_B} \ar[d]_{g} &
 \St(\Phi, B) \ar[r]^{\pi_B} \ar[d]_{h} & \St(\Phi, k[X]) \ar[d]_{i} \\
 C(\Phi, R, I) \ar[r]^{j_R}         & \St(\Phi, R, I) \ar[r]^{\mu_R} \ar@{-->}[ur]^{t}&
 \St(\Phi, R) \ar[r]^{\pi_R}        & \St(\Phi, k[X, X^{-1}]).
}\end{equation*}
Let $a$ be an element of $\Ker(h)$. Since $i$ is injective by~\cref{lem:field-case}, the element $a$ also lies in $\Ker(\pi_B)$ and hence comes from some $b \in \St(\Phi, B, I)$ via $\mu_B$. Since $g(b) \in \Ker(\mu_R)$ there exists some $c \in C(\Phi, R, I)$ such that $j_R(c) = g(b)$. By our assumption, $f$ is surjective, therefore
 $c = f(d)$ for some $d \in C(\Phi, R, I)$ and we obtain the required assertion:
 \[ 1 = \mu_Bj_B(d) = tgj_B(d) = tj_Rf(d) =t(g(b)) = \mu_B(b) = a. \qedhere \]

@@

\begin{theorem} Suppose $A$ is a local ring of characteristic $\neq 2$, then for $\ell \geq 7$
  the canonical map $\St(\rD_\ell, B) \to \St(\rD_\ell, R)$ is injective.
\end{theorem}
\begin{proof}
 In view of the above proposition and~\cref{lem:lemma32} we are left to verify that the map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective. Consider the following commutative diagram, whose rows are given by \cref{characterization}:
 \begin{equation*}\xymatrix{
  \K_3(\rD_\ell, B) \ar[r] \ar[d] & \K_3(\rD_\ell, k[X]) \ar[d] \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
  \K_3(\rD_\ell, R) \ar[r]        & \K_3(\rD_\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
 From~\cref{lem:bass-ft,lem:summary} we obtain that $\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k)$.
 In view of~\cref{lem:Panin-stability} the central vertical map is isomorphic to the canonical map
 $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$ and therefore is surjective. The requred surjectivity of $f$ now follows.
\end{proof}
